---
title: "About the c2d file format"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{About the c2d file format}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

WARNING: 🚧 _This is work in progress, please read with care._ 🚧

```{r setup}
library(c2dParseR)
```

The goal of this text is to provide a brief overview of the `.c2d` file format from [Cyclus2 ergometers by RBM elektronik-automation GmbH](https://www.cyclus2.com/en/).
The text gives an overview what kind of data is contained and how it is organized.
The intended audience are users who want to extract their research- or training-relevant data from `.c2d` files for analysis in R.
Any contained data for internal use by the Cyclus2 software is not of interest and thus omitted here.

## Disclaimer

This text is neither an authorative nor comprehensive reference for the `.c2d` file format.
The descriptions are based on the authors' understanding, some of which was verified with RBM, but some of which is based on guesses and may be incomplete or inaccurate.

## Structure overview

Technically, `.c2d` files are gzip-compressed XML text files with Cyclus2-specific content.
Let's take a tour through an "idealized" `.c2d` file to get an overview of the structure and content of the data.
The rough XML structure inside a `.c2d` file is as follows:
Inside the `C2D` node, there are 3 main nodes, `System`, `Program`, and `Training`:

- `System` contains information about the athlete, the bike settings, and the environment.
- `Program` contains information about the Cyclus2 program.
- `Training` contains the data of the training session, including the date, the header, and the row data for each time point.

```xml
<?xml version="1.0" encoding="UTF-8"?>
<BODY_50>
  <C2D>
    <System>
      <Athlete>[...]</Athlete>
      <CycleData>[...]</CycleData>
      <Field>[...]</Field>
      <VirtDerailleur>[...]</VirtDerailleur>
      <SysName>somename, somefirstname</SysName>
    </System>
    <Program>
      [...]
    </Program>
    <Training>
      <Date>somedate,sometime</Date>
      <Header>someheaders</Header>
      <Run>
        <RRInt>somerrint</RRInt>
        <Count>somecount</Count>
        <Rows>
          [...]
        </Rows>
        <Delay>somedelay</Delay>
      </Run>
    </Training>
  </C2D>
</BODY_50>
```

## System content

The `System` node contains information about the athlete, the bike settings, and the environment.

### Athlete content

The `Athlete` node contains information about the athlete, see manual, section "6.1. Athlete data".
All are user-entered in the Cyclus2 menu item _System_ \> _Athlete Settings_.

Here an example for the `Athlete` node.

```xml
<Athlete>
  <Firstname>somefirstname</Firstname>
  <Name>somename</Name>
  <Sex>1</Sex>
  <DoB>05/25/1974 12:00:00 AM</DoB>
  <Area>0.44</Area>
  <Weight>69.0</Weight>
  <Height>1.840</Height>
  <Cw>0.715</Cw>
  <tp>+etrxH[...]</tp>
</Athlete>
```

Here the meanings of the fields:

- `Firstname`: String for the first name.
- `Name`: String for the family name.
- `Sex`: Numeric code for the sex; value `1` for "male"? FIXME: Check for other possible values.
- `DoB`: Date of birth, in the format "MM/DD/YYYY 12:00:00 AM".
- `Area`: Drag Area in square meters (m²), see manual.
- `Weight`: Weight in kilograms (kg).
- `Height`: Height in meters (m).
- `Cw`: Drag coefficient, see manual.
- `tp`: Obsolete; was used for encrypted credentials for online training platforms such as by company TrainingPeaks, LLC.

### CycleData content

The `CycleData` node contains information about the bike settings.
These bike parameters are user-entered in the Cyclus2 menu item _System_ \> _Bike Settings_.
(Note that their units may be different, e.g., Crank Length is entered in millimeters, but saved in meters inside the `crank` field).
The bike parameters must accurately correspond to the actual, physical bike used for ergometry, because they enter the calculations for the values pedal force, cadence, speed, inclination and cycled distance.

Here an example for `CycleData` node; all contain numeric values.

```xml
<CycleData>
  <crank>0.1725</crank>
  <perim>2.113</perim>
  <weight>8</weight>
  <front>52</front>
  <rear>11</rear>
  <vfront>53</vfront>
  <vrear>17</vrear>
</CycleData>
```

Here their meanings, though some still seem slightly unclear.

- `crank`: The crank length, in meters (m).
- `perim`: The wheel perimeter/circumference, in meters (m).
- `weight`: The bike weight, in kilograms (kg).
- `front`: The number of teeth on the physical/real front chain ring.
- `rear`: The number of teeth on the physical/real rear sprocket.
- `vfront`: FIXME: Unclear? Probably the number of teeth on the virtual front chainring, for virtual gear settings.
- `vrear`: FIXME: Unclear? Probably the number of teeth on the virtual rear sprocket, for virtual gear settings.

### Field content

The `Field` node contains information about the environment, see manual, section "6.3. Environmental conditions".
These are user-entered in the Cyclus2 menu item _System_ \> _External Conditions_.

Here an example for the `Field` node; both contain numeric values.

```xml
<Field>
  <AirDens>1.202</AirDens>
  <FricIdx>1</FricIdx>
</Field>
```

Here their meanings:

- `AirDens`: Air density, in kilograms per cubic meter (kg/m³), see manual.
- `FricIdx`: Friction index, important for the track simulation, see manual.

### TODO: VirtDerailleur content

TODO: Describe!

## Program content (omitted)

The content of the `Program` node is currently left undescribed.
It contains internal information about the Cyclus2 program, but does not appear to be of primary interest for general data analysis.

## TODO: Training content

TODO: Describe!
