---
title: "About the c2d file format"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{about-c2d-format}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

WARNING: 🚧 _This is work in progress, please read with care._ 🚧

```{r setup}
library(c2dParseR)
```

The goal of this text is to provide a brief overview of the `.c2d` file format from [Cyclus2 ergometers by RBM elektronik-automation GmbH](https://www.cyclus2.com/en/).
The text gives an overview what kind of data is contained and how it is organized.
The intended audience are users who want to extract their research- or training-relevant data from `.c2d` files for analysis in R.
Any contained data for internal use by the Cyclus2 software is not of interest and thus omitted here.

## Disclaimer

IMPORTANT:
The text is neither an authorative nor comprehensive reference for the `.c2d` file format.
The descriptions are based on the authors' understanding, some of which was verified with RBM, but some of which is based on guesses and may be incomplete or inaccurate.

## Structure overview

Technically, `.c2d` files are gzipped XML text files with custom content.
Let's walk through an "idealized" `.c2d` file, showing the structure of the XML content.
The rough XML structure inside a `.c2d` file is as follows:
Inside the `C2D` node, there are 3 main nodes, `System`, `Program`, and `Training`:

- `System` contains information about the athlete, the bike settings, and the environment.
- `Program` contains information about the Cyclus2 program.
- `Training` contains the data of the training session, including the date, the header, and the row data for each time point.

```xml
<?xml version="1.0" encoding="UTF-8"?>
<BODY_50>
  <C2D>
    <System>
      <Athlete>[...]</Athlete>
      <CycleData>[...]</CycleData>
      <Field>[...]</Field>
      <VirtDerailleur>[...]</VirtDerailleur>
      <SysName>somename, somefirstname</SysName>
    </System>
    <Program>
      [...]
    </Program>
    <Training>
      <Date>somedate,sometime</Date>
      <Header>someheaders</Header>
      <Run>
        <RRInt>somerrint</RRInt>
        <Count>somecount</Count>
        <Rows>
          [...]
        </Rows>
        <Delay>somedelay</Delay>
      </Run>
    </Training>
  </C2D>
</BODY_50>
```
